Accessible, realistic genome simulation with selection using \stdpopsim

### 2025-06-12 review-responses.tex

  \begin{point}{\revref} % \&75
    How do you deal with overlapping annotations in a gff?
  \end{point}

  \reply{
    TODO. \revref
  }

  \begin{point}{}
    The inference of the deleterious DFE with GRAPES is quite inaccurate. 
    Is it possible that there is a factor of 2 error involved?
  \end{point}

  \reply{
As noted in the paper, conversion of selection inferences between tools is challenging.
We thus consulted with the developers of \polydfe and \grapes to confirm our conversions.
(The developer of \dadicli is an author on this paper.)
The inaccuracy of \grapes is a surprise to us.
Note that \grapes is also less accurate than \polydfe and \dadicli in inferring the shape of the distribution (Fig.~\ref{fig:homsap-dfe.ooa}B), which suggests the inaccuracy is not simply a scaling issue.
  }

### 2025-06-12 analysis2_manuscript.tex
Response to factor of two review comment. Also updates to methods to clarify scaling.

    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on species' publicly available functional genomic elements
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
    occurring outside of these regions do not affect fitness.
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for five species in the catalog:
    \textit{Arabidopsis thaliana} \citep{huber2018gene},
    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining,zhen2021greater},
    humans \citep{huber2017determining,kim2017inference,kyriazis2023using,castellano2019compmarison,castellano2019compmarison,rodrigues2024shared},
    \textit{Mus musculus} \citep{booker2021selective},
    and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of species, these DFEs can be applied to other species
    in the catalog if and when it is suitable,
    namely, when a species of interest lacks genomic resources \revpoint{1}{8}.
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2023using}
    for ``general'' use across mammals (lacking more specific information).
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have
    a published DFE included in the catalog.

@@

    % ADK: I've added a section below that describes the DFE inference methods used in this study.
    To facilitate direct comparison between inference methods,
    we scale each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
    %IG: expanded a bit to justify each scaling decision
    \grapes outputs the mean and shape of the inferred distribution of $4N_es$.
    We thus scale the mean by a factor of  $\frac{\mu}{\theta}$,
    where $\mu$ is the assumed mutation rate and
    $\theta=4N_{e}\mu$ is the population-scaled mutation rate, as estimated by \grapes.
    \polydfe is similarly scaled by $\frac{2\mu}{\theta}$,
    reflecting the fact that it defines the selection coefficient such that a homozygote has fitness $1+ 2s$.
    \dadicli infers a gamma distribution for $N_es$ and parameterizes this distribution using the shape ($\alpha$) and rate ($\nu$).
    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
    Thus, the mean value of $s$ is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$,
    where $L$ is the total number of sites available for the analysis,
    and $p$ is the proportion of sites under selection.
    Complete details for the DFE inference analysis can be found in the
    \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.

>>>>

    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on species' publicly available functional genomic elements
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
    occurring outside of these regions do not affect fitness.
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for five species in the catalog:
    \textit{Arabidopsis thaliana} \citep{huber2018gene},
    \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining,zhen2021greater},
    humans \citep{huber2017determining,kim2017inference,kyriazis2023using,castellano2019compmarison,castellano2019compmarison,rodrigues2024shared},
    \textit{Mus musculus} \citep{booker2021selective},
    and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of species, these DFEs can be applied to other species
    in the catalog if and when it is suitable.
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Indeed, the DFE labeled \texttt{HomSap/Mixed\_K23} is recommended by \citet{kyriazis2023using}
    for ``general'' use across mammals (lacking more specific information).
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have
    a published DFE included in the catalog.

@@

    % ADK: I've added a section below that describes the DFE inference methods used in this study.
    To facilitate direct comparison between inference methods,
    we scale each program's output to obtain a selection coefficient ($s$) consistent with the one used by \slim.
    %IG: expanded a bit to justify each scaling decision
    \grapes outputs the mean and shape of the inferred distribution of $4N_es$.
    We thus scale the mean by a factor of  $\frac{\mu}{\theta}$,
    where $\mu$ is the assumed mutation rate and
    $\theta=4N_{e}\mu$ is the population-scaled mutation rate, as estimated by \grapes.
    \polydfe is similarly scaled by $\frac{2\mu}{\theta}$,
    reflecting the fact that it defines the selection coefficient such that a homozygote has fitness $1+ 2s$.
    \dadicli defines the selection coefficient similarly to \polydfe and infers a gamma distribution for $2 N_e s$, parameterized by the shape ($\alpha$) and rate ($\nu$).
    It also infers the total population-scaled mutation rate for selected sites over the considered region $\theta = 4 N_e \mu\, L\, p$,
    where $L$ is the total number of sites available for the analysis,
    and $p$ is the proportion of sites under selection.
    %IG: we're still missing justification for the scaling by L and p, but we can leave this for revision
    Thus, the mean value of $s$ in \slim terms is obtained by  $\frac{4\mu}{\theta} \alpha \nu L p$.
    Complete details for the DFE inference analysis can be found in the
    \snakemake workflow at \url{https://github.com/popsim-consortium/analysis2/blob/main/workflows/dfe.snake}.
